Title here

### 2016-12-05 portfolio/text/ms.Rmd
Start developing result section

The state and federal fisheries of Alaska, provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability for those whose livelihood depends on the extraction of natural
resources. Fisheries in Alaska are among the highest-valued fisheries in the
world with 2.7 billion kg landed biomass valued at nearly 2 billion USD annually
[@nmfs2016]. The industry supports a range of fishing fleets ranging from small,
open, single-operator vessels to multi-million-dollar state-of-the-art ships
with large crews. Like much of the rest of the world [@hilborn2003a], the days
of open, unregulated fisheries in most developed countries have passed ---
fishers must own a license to participate in virtually all fisheries. And while
such regulations have proven helpful in reducing or eliminating overfishing,
they limit the available strategies to reduce income variability. For example,
most permits limit fishing to a single species, diversifying by purchasing
additional fishing permits can be expensive (e.g. ~\$XX for king crab), and some
fisheries limit the entry of new participants through individual fishing quotas
(IFQs) (REF).

@@

**Fig. \ref{fig:ts} - limited entry + movement towards salmon and specialization
in Alaska**

- move towards specialization, salmon fig 1D (overall declining trend disappears if you exclude salmon)
- halibut most often combined with other permits (exclude permit holders that hold any halibut and about ninety percent of the remaining permit holders hold only one permit) fig 1c
- Salmon the most often on its own (exclude salmon and the proportioned single permit decreases to about 70%) fig 1c
- majority are single permit holders (70-90%) fig 1C, but the single permit holders make up only about half of the overall revenue (and surpassed half for the first time in 2013 and stayed that way) fig 1B

@@

**Fig. \ref{fig:across} - across strategies**

- clear exponential decline in standardized variability with strategies that are inherently more species diverse
- not a clear effect of days fished at this level
- fishing permits with higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity);

**Fig. \ref{fig:spokes} - add permits**

- nearly universal reduction in variability (Fig. S\ref{fig:spoke-all})
- e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR]
lower expected standard deviation
for halibut plus another permit vs. single halibut permits);
- usually associated with higher median revenue
- but can be very expensive (provide some examples from website)

>>>>

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability for those whose livelihood depends on the extraction of natural
resources. Fisheries in Alaska are among the highest-valued fisheries in the
world with 2.7 billion kg landed biomass valued at nearly 2 billion USD annually
[@nmfs2016]. The industry supports a range of fishing fleets ranging from small,
open, single-operator vessels to multi-million-dollar state-of-the-art ships
with large crews. Like much of the rest of the world [@hilborn2003a], the days
of open, unregulated fisheries in most developed countries have passed ---
fishers must own a license to participate in virtually all fisheries. And while
such regulations have proven helpful in reducing or eliminating overfishing,
they limit the available strategies to reduce income variability. For example,
most permits limit fishing to a single species, diversifying by purchasing
additional fishing permits can be expensive (e.g. ~\$XX for king crab), and some
fisheries limit the entry of new participants through individual fishing quotas
(IFQs) (REF).

@@

Trends in fishing revenue and diversification over time for individuals fishing
commercially in Alaska (Fig. \ref{fig:ts}A) reveal some movement toward
specialization, and in particular specialization on salmon. The majority
(70--90%) of fishers hold single permits, with a slight increase in single
permit holders since around 2000 (Fig. \ref{fig:ts}C). This overall trend is
most sensitive to individuals with halibut or salmon licenses --- halibut permit
owners most often have multiple permits whereas salmon permit holders tend to be
single permit holders more than individuals with permits for other fisheries
(Fig. \ref{fig:ts}C). Although the majority of individuals hold single permits,
multiple permit holders make approximately half of the overall revenue (Fig.
\ref{fig:ts}B). Switching focus from permits owned to species landed, mean
fished species diversity declined somewhat from 1980--2000 but has since
remained steady (Fig. \ref{fig:ts}D). However, this trend appears to be driven
by an increase in salmon specialists in the 1980s and 1990s --- the trend
disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).

@@

We observed a clear exponential decrease in revenue variability for strategies
(groups of permits) with higher mean fished species diversity (Fig.
\ref{fig:across}). Strategies with one unit higher mean species diversity had
on average
`r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) 
lower standardized variability. This pattern looks qualitatively similar with or
without individuals fishing salmon. The gillnet fishery for herring roe and the
king crab fishery had the highest estimated revenue variability. Most of the
strategies with the lowest estimated revenue variability combined halibut and
salmon permits and had high levels of fished species diversity (Fig.
\ref{fig:across}).

Comparing single permit holders to other individuals holding that same permit
plus one or more other permits illustrates the expected reduction in revenue
variability from perching additional permits (Fig. \ref{fig:spokes},
Fig. S\ref{fig:spoke-all}). For example, individuals fishing on a halibut longline
permit plus an additional permit always had lower expected revenue variability
compared to individuals fishing only a halibut longline permit 
(median reduction of `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet permit
substantially reduced expected revenue variability. Adding permits also usually
resulted in higher expected revenue, although the permits themselves represent
a significant expense (e.g. XX).
